\documentclass{article}
\pdfoutput=1 


% \usepackage[version=3.2]{biblatex}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024


% ready for submission
\usepackage[nonatbib,final]{neurips_2024}


% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
% \usepackage[preprint]{neurips_2024}


% to compile a camera-ready version, add the [final] option, e.g.:
    % \usepackage[final]{neurips_2024}


% to avoid loading the natbib package, add option nonatbib:
   % \usepackage[nonatbib]{neurips_2024}


\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
% \usepackage{xcolor}         % colors
\usepackage{soul}
\usepackage{appendix}

\input{preamble}

\title{What Representational Similarity Measures Imply about Decodable Information}


% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.


\author{%
  Sarah E. Harvey\\
  Center for Computational Neuroscience\\
  Flatiron Institute, New York, NY\\
  \texttt{sharvey@flatironinstitute.org}
  \And
  David Lipshutz\\
  Center for Computational Neuroscience\\
  Flatiron Institute, New York, NY\\
  \texttt{dlipshutz@flatironinstitute.org}
  \And
  Alex H. Williams\\
  Center for Neural Science, Center for Computational Neuroscience\\
  New York University, Flatiron Institute, New York, NY\\
  \texttt{alex.h.williams@nyu.edu}
}

\begin{document}


\maketitle


\begin{abstract}
Neural responses encode information that is useful for a variety of downstream tasks.
A common approach to understand these systems is to build regression models or ``decoders'' that reconstruct features of the stimulus from neural responses.
Popular neural network similarity measures like centered kernel alignment (CKA), canonical correlation analysis (CCA), and Procrustes shape distance, do not explicitly leverage this perspective and instead highlight geometric invariances to orthogonal or affine transformations when comparing representations.
Here, we show that many of these measures can, in fact, be equivalently motivated from a decoding perspective.
Specifically, measures like CKA and CCA quantify the average alignment between optimal linear readouts across a distribution of decoding tasks. 
We also show that the Procrustes shape distance upper bounds the distance between optimal linear readouts and that the converse holds for representations with low participation ratio.
Overall, our work demonstrates a tight link between the geometry of neural representations and the ability to linearly decode information.
This perspective suggests new ways of measuring similarity between neural systems and also provides novel, unifying interpretations of existing measures. 
\end{abstract}


\input{01-intro}
\input{02-setup_revision}
\input{03-cka-gulp_revision}
\input{04-shapes}
\input{06-discussion}

\printbibliography
\clearpage

\begin{appendices}
\input{999-appendix}
\input{05-generalization-revision}
\end{appendices}

\end{document}
